---
title: "Determinants of Maritime Piracy"
author: "Bj?ern Boening, Cody Koebnick, Laurence Hendry"
date: "12. November 2015"
output:
  html_document: default
  pdf_document:
    number_section: true
---
# Introduction

Maritime piracy has been in the media, especially with attention directed to the Horn of Africa. ATALANTA, the European Union's naval mission, patrolled along the coast of Somalia to combat maritime piracy and protect trading vessels against attacks. As it seems, fading media attention does not presuppose a reduced danger of maritime piracy globally.

Global shipping routes are highly important for trade. Piracy attacks are a potential threat for crew and cargo on the ship. The cost intensive deployment of international naval forces in Somalia shows how serious countries take the threat whose trade is affected. Interestingly not all piracy attacks are successful, and the ratio varies from country to country and over time. So what drives piracy attacks, why and when are they successful?

# Research Design

The idea for this paper is to research a model which can explain piracy under which circumstances attacks are successful. 

Does the number of attacks decrease the likelihood of attacks being successful?

We expect that several variables can 

# Data 

The International Maritime Bureau (IMB) collects since 1992 all reported piracy attack globally. Since then it publishes annually an overview of all attacks that happened in a year. These annual reports provided by the IMB contain detailed information about every incident, which allows for further analysis of distinct types of piracy attacks, for instance successful attacks v. attempted attacks.

The annual reports were scraped with text analysis tools. Our team received a "ready to use" dataset from a research project from the university of Tennessee, including all global piracy attacks from 1994 to 2014.


# Additional data gathering

The original dataset contains the attacks that were reported by the victims of piracy. In our r 

hhhh
gi
```{r, echo=FALSE}
try()
try(setwd("/Users/codykoebnick/Downloads/Data Set"))
shipping <- read.csv("total4 naughty2.csv", header = TRUE, sep = ",", stringsAsFactors = FALSE, na.strings = c("", "NA"))

#renaming and dropping some of our columns 
names(shipping)[1] <- 'country'
names(shipping)[2] <- 'year'
shipping$X1 = NULL
names(shipping)[3] <- 'coast/Area ratio (m/km2)'
names(shipping)[4] <- 'GDP per cap'
names(shipping)[5] <- 'attacks/Year'
names(shipping)[6] <- 'successful Attacks/Year'
names(shipping)[7] <- 'success Ratio'

#######################
#Descriptive Statistics
#######################

#Histogram of attack frequnecy
hist(shipping$`attacks/Year`)
```



# Data gathering

We 
```{r}
summary(cars)
```

You can also embed plots, for example:

```{r, echo=FALSE}
plot(cars)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
