---
title: "Determinants of Maritime Piracy"
author: "Bj?ern Boening, Cody Koebnick, Laurence Hendry"
date: "12. November 2015"
output:
  html_document: default
  pdf_document:
    number_section: true
---
```{r, echo=FALSE, message=FALSE}
library(base)
library(rio) # swiss army knife for imports
library(plyr) # count occurences
library(dplyr) # data wrangling
library(tidyr) # data wrangling
library(ggplot2) # nice plots
library(stargazer) # nicer regression output which looks like a real publication
library(car) # scatterplots 
library(httr) # scraping from http sites
library(XML) # Tool for generating XML file
library(WDI) # Scraping Data from the World Bank 
library(countrycode) # provides world bank country codes 
library(gplots)
library(plm)
```


# I Introduction

Maritime piracy has been in the media, especially with attention directed to the Horn of Africa. ATALANTA, the European Union's naval mission, patrolled along the coast of Somalia to combat maritime piracy and protect trading vessels against attacks. As it seems, fading media attention does not presuppose a reduced danger of maritime piracy globally.

Global shipping routes are highly important for trade. Piracy attacks are a potential threat for crew and cargo on the ship. The cost intensive deployment of international naval forces in Somalia shows how serious countries take the threat whose trade is affected. Interestingly not all piracy attacks are successful, and the ratio varies from country to country and over time. So what drives piracy attacks, why and when are they successful?

# II Research Design

The idea for this paper is to research a model which can explain piracy under which circumstances attacks are successful. 

Does the number of attacks decrease the likelihood of attacks being successful?

The dependent variable is the success rate of piracy attacks, calculated by the number of successful attacks divided by the total number of attacks. We expect that mainly the total number of attacks has an impact on this ratio. The fact that the dependent variable actually consists of our key independent variable is dangerous. However, we think that there must be a visible learning effect, either from law enforcement bodies, the shipping crew, or the pirates. So far this was the only feasible way we could have a look at this effect. 

Furthermore, additional exogenous variables will be included that in theory should have an impact on the inspected success rate. GDP (per capita per year) as a mirror for the economic incentives to conduct piracy is expected to influence the success rate over time. Likewise, a country's ratio of coast line to its land area should be a good demarker, whether piracy attacks happen more often. 

1. After a certain amount of piracy attacks, the success rate has a peek and will decrease.
2. The higher the GDP, the less incentive there is to conduct an attack and only less skilled pirates will make attempts, hence the success rate decreases.
3. The higher the coast-land ratio is, the more people decide to conduct attacks and skilled pirates emerge, hence the success rate increases.


# III Preparation of Dataset

### Principal Gathering

Since 1992, the International Maritime Bureau (IMB) has collected all reported piracy attack globally. Since then it publishes annually an overview of all attacks that happened in a year. These annual reports provided by the IMB contain detailed information about every incident, which allows for further analysis of distinct types of piracy attacks, for instance successful attacks v. attempted attacks.

The annual reports were scraped with text analysis tools. Our team received a "ready to use" dataset from a research project from the university of Tennessee, including all global piracy attacks from 1994 to 2014.

### Additional Data Gathering: World Bank and Wikipedia

The original dataset contains the attacks that were reported by the victims of piracy. As an additional variable, relevant to our field of piracy investigation and patrols, we were intrigued by the relative effect that a longer coast length of a country has on the level of attacks that country suffers. To address this question we  parsed a table titled "List of countries by length of coastline" from a Wikipedia page that had, in turn, used information from the CIA World Factbook. We then merged this coastline data with our existing dataframe using a 'right outer join'. 

Of critical interest to us were the respective 'Coast/Area' ratios (measured in km of coast length to km of square land) that serves as an insightful control for our country dependent variable. 

### A Work Around with Excel

# IV Visualizing and Examining the Data

Below is a histogram showing the frequency of attacks sorted by number of attacks per year.  In other words, we can clearly see that the vast majority of years had 1-20 attacks reported.  The frequency then goes gradually within the 20-40 range and then again in the 40-60 range. We see a slight surge in number of attacks per year within the 80-100 range, this is due to highly pirated coastal areas such as Bangladesh and the philippines. 

```{r, echo=FALSE}
try()
try(setwd("/Users/codykoebnick/Documents/MaritimePiracy"))
shipping <- read.csv("shippingraw.csv", header = TRUE, sep = ",", stringsAsFactors = FALSE, na.strings = c("", "NA"))

#renaming and dropping some of our columns 
names(shipping)[1] <- 'country'
names(shipping)[2] <- 'year'
shipping$X1 = NULL
names(shipping)[3] <- 'coast/Area ratio (m/km2)'
names(shipping)[4] <- 'GDP per cap'
names(shipping)[5] <- 'attacks/Year'
names(shipping)[6] <- 'successful Attacks/Year'
names(shipping)[7] <- 'success Ratio'

#######################
#Descriptive Statistics
#######################

#Histogram of attack frequnecy
hist(shipping$`attacks/Year`)
```

Similarily, the sucess of attacks is distributed in a very similar fashion to the number of attacks.  However, our study seeks to see if there is indeed a statistically significant relationship between the number of attacks and attack success.   

```{r, echo=FALSE}
#Histogram of successful attacks per year
hist(shipping$`successful Attacks/Year`)
```

Below, a plot displays the relationship between attack success and the GDP per capita of the closest coastal state.  

```{r, echo=FALSE}
#histogram of attack success over GDP per cap
plot(shipping$`GDP per cap`, shipping$`success Ratio`)
```

Here we have seperate coplots depicting the success ratio for attacks by year, sorted by the closest coastal state. 

```{r, echo=FALSE}
#exploring the data 
coplot(shipping$`success Ratio` ~ year|country, type="l", data=shipping) # Lines
```

When we examine the heterogeneity across countries it appears that a pirates change of a successful attack is dependent on the closest coastal state.  For instance, the changes of a successful attack in the Philippines is drastically lower than the odds of a successful attack in Bangladesh.  This also holds true when considering confidence intervals which are also depicted.  

```{r, echo=FALSE}
# Fixed effects: Heterogeneity across countries (or entities)
# interestingly, attacks in Philipines have the lowest success ratio and attacks in Brazil have the highest.
# plotmeans draw a 95% confidence interval around the means
plotmeans(shipping$`success Ratio` ~ country, main="Heterogeineity across countries", data=shipping)
```

When we examine the heterogeneity across years the means seem to stay within the .65 - .85 range.  However, due to extremely large confidence intervals, time does not appear to be statistically significant.  

```{r, echo=FALSE}
# Fixed effects: Heterogeneity across years
# plotmeans draw a 95% confidence interval around the means
plotmeans(shipping$`success Ratio` ~ year, main="Heterogeineity across years", data=shipping)
```

Below is our first OLS regrssion.  Although OLS regrssion does not consider heterogeneity across groups or time, an OLS regrssion can still prove useful for gathering initial insight into the relationship of our variables.

For instance, here we see that attacks per year does not a statistically significant effect with a p-value of >.1.  However, an interesting point worth noting would be that negative coefficient of attacks per year. 

```{r}
# OLS Regression
# Regular OLS regression does not consider heterogeneity across groups or time.
# In this simple model, the number of attacks has a slightly negative relationship with attack success, however it is not stat. sig. 
ols <-lm(shipping$`success Ratio` ~ shipping$`attacks/Year`, data=shipping)
summary(ols)
```
The below plot shows that after attacks in a certain country reach a threshold, approximately 40, their attack success ratio is steadily above .6.
```{r, echo=FALSE}
# The below plot shows that after attacks in a certain country reach a threshold, approximately 40, their attack success ratio is steadily above .6
yhat <- ols$fitted
plot(shipping$`attacks/Year`, shipping$`success Ratio`, pch=19, xlab="x1", ylab="y")
abline(lm(shipping$`success Ratio`~shipping$`attacks/Year`),lwd=3, col="red")
```

In our model's first fixed effects regression, attacks per year becomes statistically significant and has a small negative coefficient of -.002.  It is important to note that the attacks per year became significant only in the fixed effects model, as opposed to the OLS.

```{r, echo=FALSE}
#attempting fixed effects model - attack success on attacks per year
fixed <- plm(shipping$`success Ratio` ~ shipping$`attacks/Year`, data=shipping, index=c("country", "year"), model="within")
summary(fixed)
```

However, once we add a variable controlling for the coast ratio of the closest coastal state, attacks per year once again becomes statistically insignificant.  

```{r, echo=FALSE}
# Increasing the complexity of our model
############################

# OLS Regression
# Regular OLS regression does not consider heterogeneity across groups or time.
# In this simple model, the number of attacks has a slightly negative relationship with attack success, however it is not stat. sig. 
# Model 2 - Success ratio on attacks/year + coast
ols2 <-lm(shipping$`success Ratio` ~ shipping$`attacks/Year` + shipping$`coast/Area ratio (m/km2)`, data=shipping)
summary(ols2)
```

Interestingly, the plot below shows that the coastal ratio of the closest coastal state most likely does have a positive, significant effect on attack success, however the Philippines is a strong outlier. 

```{r, echo=FALSE}
yhat2 <- ols$fitted
plot(shipping$`coast/Area ratio (m/km2)`, shipping$`success Ratio`, pch=19, xlab="x1", ylab="y")
abline(lm(shipping$`success Ratio`~shipping$`coast/Area ratio (m/km2)`),lwd=3, col="red")
```

When we add GDP per capita nothing is significant. 

```{r, echo=FALSE}
### Model 3 - Sucess rati on a
# attacks/year + coast + GDP
ols3 <-lm(shipping$`success Ratio` ~ shipping$`attacks/Year` + shipping$`coast/Area ratio (m/km2)` + shipping$`GDP per cap`, data=shipping)
summary(ols3)
```

A graph of the above regression

```{r, echo=FALSE}
yhat3 <- ols$fitted 
plot(shipping$`GDP per cap`, shipping$`success Ratio`, pch=19, xlab="x1", ylab="y")
abline(lm(shipping$`success Ratio`~shipping$`GDP per cap`),lwd=3, col="red")
```

Number of attacks is significant again.
```{r, echo=FALSE}
# FE model 2, Sucess Ratio on Attacks/Year + GDP per cap 
#attempting fixed effects model
fixed2 <- plm(shipping$`success Ratio` ~ shipping$`attacks/Year` + shipping$`GDP per cap`, data=shipping, index=c("country", "year"), model="within")
summary(fixed2)
```

Here are diagnostic regressions

```{r}
#####################
# Regression Diagnostics
######################

#Assessing homoscedasticity (we have met the constant variance assumption if p < 1.95)
ncvTest(ols2)
```

```{r}
#Assessing multicollinearity
vif(ols2)
```

```{r}
#Assessing outliers
outlierTest(ols2)
```






