---
title: "Determinants of Maritime Piracy"
author: "Bjöern Boening, Cody Koebnick, Laurence Hendry"
date: "12. November 2015"
output:
  html_document: default
  pdf_document:
    number_section: true
---
# Introduction

Maritime piracy has been in the media, especially with attention directed to the Horn of Africa. ATALANTA, the European Union's naval mission, patrolled along the coast of Somalia to combat maritime piracy and protect trading vessels against attacks. As it seems, fading media attention does not presuppose a reduced danger of maritime piracy globally.

Global shipping routes are highly important for trade. Piracy attacks are a potential threat for crew and cargo on the ship. The cost intensive deployment of international naval forces in Somalia shows how serious countries take the threat whose trade is affected.

# Research Question

The idea for this paper is to research a model which can explain piracy attacks and also examine the development of piracy hotspots around the world over time. 

What determines maritime piracy?

```{r}
summary(cars)
```

You can also embed plots, for example:

```{r, echo=FALSE}
plot(cars)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
