---
title: "Determinants of Maritime Piracy"
author: "Bjoern Boening, Cody Koebnick, Laurence Hendry"
date: "10 December 2015"
output:
  html_document: default
  pdf_document:
    number_section: yes
    toc: yes
    fig_caption: yes
---
```{r, echo=FALSE, message=FALSE, warning=FALSE}
#helpful shortcuts for editing what is includeded in Knitr: #http://kbroman.org/knitr_knutshell/pages/Rmarkdown.html
library(base)
library(rio) # swiss army knife for imports
library(plyr) # count occurences
library(dplyr) # data wrangling
library(tidyr) # data wrangling
library(ggplot2) # nice plots
library(stargazer) # nicer regression output which looks like a real publication
library(car) # scatterplots 
library(httr) # scraping from http sites
library(XML) # Tool for generating XML file
library(WDI) # Scraping Data from the World Bank
library(countrycode) # provides world bank country codes 
library(gplots)
library(plm)
library(knitr)
library(threejs)
library(maptools)
library(maps)
library(Amelia)

shipping <- read.csv("Data/shippingraw.csv", header = FALSE, sep = ",", stringsAsFactors = FALSE, na.strings = c("", "NA"))

#renaming and dropping some of our columns 
names(shipping)[1] <- 'country'
names(shipping)[2] <- 'year'
shipping$X1 = NULL
names(shipping)[3] <- 'coast/Area ratio (m/km2)'
names(shipping)[4] <- 'GDP per cap'
names(shipping)[5] <- 'attacks/Year'
names(shipping)[6] <- 'successful Attacks/Year'
names(shipping)[7] <- 'success Ratio'
```


# I Introduction

Maritime piracy has been in the media, especially with attention directed to the Horn of Africa. ATALANTA, the European Union's naval mission, patrolled along the coast of Somalia to combat maritime piracy and protect trading vessels against attacks. As it seems, fading media attention does not presuppose a reduced danger of maritime piracy globally.

Global shipping routes are highly important for trade. Piracy attacks are a potential threat for crew and cargo on the ship. The cost intensive deployment of international naval forces in Somalia shows how serious countries take the threat whose trade is affected. Interestingly not all piracy attacks are successful, and the ratio varies from country to country and over time. So what drives piracy attacks, why and when are they successful?

# II Research Design

The idea for this paper is to research a model which can explain piracy under which circumstances attacks are successful. 

Does the number of attacks decrease the likelihood of attacks being successful?

The dependent variable is the success rate of piracy attacks, calculated by the number of successful attacks divided by the total number of attacks. We expect that mainly the total number of attacks has an impact on this ratio. The fact that the dependent variable actually consists of our key independent variable is dangerous. However, we think that there must be a visible learning effect, either from law enforcement bodies, the shipping crew, or the pirates. So far this was the only feasible way we could have a look at this effect. 

Furthermore, additional exogenous variables will be included that in theory should have an impact on the inspected success rate. GDP (per capita per year) as a mirror for the economic incentives to conduct piracy is expected to influence the success rate over time. Likewise, a country's ratio of coast line to its land area should be a good demarker, whether piracy attacks happen more often. 

1. After a certain amount of piracy attacks, the success rate has a peek and will decrease.
2. The higher the GDP, the less incentive there is to conduct an attack and only less skilled pirates will make attempts, hence the success rate decreases.
3. The higher the coast-land ratio is, the more people decide to conduct attacks and skilled pirates emerge, hence the success rate increases.


# III Preparation of Dataset

### Principal Gathering

Since 1992, the International Maritime Bureau (IMB) has collected all reported piracy attack globally. Since then it publishes annually an overview of all attacks that happened in a year. These annual reports provided by the IMB contain detailed information about every incident, which allows for further analysis of distinct types of piracy attacks, for instance successful attacks v. attempted attacks.

The annual reports were scraped with text analysis tools. Our team received a "ready to use" dataset from a research project from the university of Tennessee, including all global piracy attacks from 1994 to 2014.

### Additional Data Gathering: World Bank and Wikipedia

The original dataset contains the attacks that were reported by the victims of piracy. As an additional variable, relevant to our field of piracy investigation and patrols, we were intrigued by the relative effect that a longer coast length of a country has on the level of attacks that country suffers. To address this question we  parsed a table titled "List of countries by length of coastline" from a Wikipedia page that had, in turn, used information from the CIA World Factbook. We then merged this coastline data with our existing dataframe using a 'right outer join'. 

Of critical interest to us were the respective 'Coast/Area' ratios (measured in km of coast length to km of square land) that serves as an insightful control for our country dependent variable. 

The information about the gross domestic product comes from the worldbank and was scraped with the WDI package for R. The scraped data comes in a country-year format, thus it comes already in a format we need to conduct our analysis.

# Cleaning

Firstly, we reduced the key dataset on piracy attacks to the 8 countries with the highest number of attacks. The remaining countries are:

1. Bangladesh
2. Brazil
3. India
4. Indonesia
5. Malaysia
6. Nigeria
7. Philippines
8. Vietnam 

We also consider high levels of activity in the Gulf of Aden and piracy around Somalia. 

Secondly, the all data needed to be merged. After the additionally gathered data was clean, we merged into the original dataset on piracy the variables for GDP and the coastline ratio. The initial plan was to reshape this dataset into panel data. However, as you will see in our R-code the reshape was not fully succsessful. We did not get rid of all first-level variables, or in other words we were unable to aggregate to the desired country_year level. We deleted all irrelevant variables, renamed the remaining ones with more intuitive names.

In the next section we explain, why we decided to continue with a new dataset produced with excel.


# A Work Around with Excel

A major challenge in the preparation of our data was the change of our units of observations for our dataset MaritimePiracyTennessee.csv from 'incidents of pirate attack' to the intended unit of 'country-year' for our prepared 'shippingraw.csv' dataset. 

After several attempts to remove duplicates and particular values in the transformation process we determined to save our intermediate progress. We then proceeded to open and transform the dataset using Microsoft Excel, using the find and replace function supported in this suite. This excursion with another suite was intentionally limited to just these two steps, with data preparation continuing in Analysis.R. 

The offending code:

**Step 1: Attempt to drop rows conditional given particular values using R**

SuccRatCtryYr <- table(shipping$year, shipping$closest_coastal_state, shipping$Incident_type_recode==1)
SuccRatCtryYr
class(SuccRatCtryYr)


**Step 2: Attempt to remove duplicate rows using R**

duplicated(total4)
newtotal4 <- total4[duplicated(total4)=='FALSE', ]

# IV Visualizing and Examining the Data
```{r, echo=FALSE}
try(setwd("/Users/laurencehendry/GitHub/MaritimePiracy"),silent=TRUE) 
globe <- read.csv("MaritimePiracyTennessee.csv", header = TRUE, sep = ";", stringsAsFactors = FALSE, na.strings = c("", "NA"))

URL <- 'https://en.wikipedia.org/wiki/List_of_countries_by_length_of_coastline'

tables <- URL %>% GET() %>%
  content(as = 'parsed') %>% 
  readHTMLTable()

names(tables)

CoastlineTable <- tables[[1]]

head(CoastlineTable)[, 1:3]

#cleaning the scrape for merge
CoastlineTable$V2 = NULL
CoastlineTable$V3 = NULL
CoastlineTable$V4 = NULL
CoastlineTable$V5 = NULL
CoastlineTable$V6 = NULL
CoastlineTable$V8 = NULL

#renaming, part of scrape
colnames(CoastlineTable)
names(CoastlineTable)
names(CoastlineTable)[1] <- 'closest_coastal_state'
names(CoastlineTable)[2] <- 'CoastRatio'

#merging
#p297 from R for Dummies
globe <- merge(globe, CoastlineTable, all.x=TRUE)


globe$longitude <- gsub(",", ".", globe$longitude)
globe$latitude <- gsub(",", ".", globe$latitude)

data(wrld_simpl, package="maptools")
data(worldMapEnv, package="maps")
```

### 3D Globes using the 'threejs' Package

Many thanks to Bryan W. Lewis and his website for the creation and help in using the [threejs package](http://bwlewis.github.io/rthreejs/). The 3D globes are full navigable, with pinch-zooming enabled for trackpads to help with examining statistical results in particular regions. 

## Global Overview: Piracy and Maritime Defence Propensity 
```{r, echo=FALSE}
######################################
# Create a Globe Viewer: Using 'CoastRatio' for arcs; zoomed out region focus; Phillipines highlighted -LH 
######################################

bgcolor <- "#000025"
earth <- tempfile(fileext=".jpg")
jpeg(earth, width=2048, height=1024, quality=100, bg=bgcolor, antialias="default")
par(mar = c(0,0,0,0), pin = c(4,2), pty = "m", xaxs = "i",
    xaxt = "n", xpd = FALSE, yaxs = "i", yaxt = "n")

col <- rep("black",length(wrld_simpl$NAME))    # Set a default country color
col[wrld_simpl$NAME=="Philippines"] <- "#FF0000"    # Highlight country


plot(wrld_simpl, col=col,          bg=bgcolor, border="green", ann=FALSE,  axes=FALSE, 
     xpd=FALSE,  xlim=c(-180,180), ylim=c(-90,90),  setParUsrBB=TRUE)

graphics.off()

globejs(img=earth, bg="black", lat=globe$latitude, long=globe$longitude, value=globe$CoastRatio, color=col, pointsize=0.5, rotationlat=0.2,     rotationlong=-3, fov=30, atmosphere=TRUE)
```
Nb. Country borders highlighted green; Philippines highlighted red; Arc length indicates height of coast-to-country ratio (small to large)

This globe provides insight on the issue of piracy 'on the high seas', by investigating whether a country's coastal length has an effect on the number and location of pirate attacks we can observe. The assumption here is that if a country has a higher coastal to country ratio (an island-state, for instance) then it should be less likely that that country is affected by maritime piracy, since there is a greater incentive to defend itself and maritime defence is therefore prioritised and greater in that country. 

H1: Most pirate attacks occur in areas where countries have smaller country-to-coast ratios, and thus are less incentivised to defend against maritime pirate  attacks. 
H0: Whether a country has a greater country-to-coast ratio has either an inverse or no effect on where pirate attacks occur. 

To observe this we have coded the arc length (shown in white) of the individual attacks that occur to show height of coastline-to-country ratios. If there is a tall arc, then the country has a proportionately higher coastal size. The expectation is therefore that fewer tall arcs should appear since these countries are actively patrolling and bordering their vulnerable coastlines. 

This is not always the case, depending on region. Our H1 remains true for the area around the Gulf of Aden and particularly in the Arabian Sea, where countries with less incentive to defend maritime areas do indeed suffer from higher attack frequency. Oman, Pakistan, India, Somalia, Kenya and Tanzania are prime examples of this pattern.

However, the results change when shifting focus to the South China Sea. Despite Indonesia and the Philippines (highlighted) having very long coastal lengths, a lot of attacks can be observed. This confirms our H0 null hypothesis, with contrary results occuring for Southeast Asian States. 


### Regional Focus: Piracy in the South China Sea
```{r, echo=FALSE}
######################################
# Create a Globe Viewer 2: Using 'Steaming.Recode' for arcs; SEAsia region focus; Indonesia highlighted -LH
######################################

bgcolor <- "#000025"
earth <- tempfile(fileext=".jpg")
jpeg(earth, width=2048, height=1024, quality=100, bg=bgcolor, antialias="default")
par(mar = c(0,0,0,0), pin = c(4,2), pty = "m", xaxs = "i",
    xaxt = "n", xpd = FALSE, yaxs = "i", yaxt = "n")

col <- rep("black",length(wrld_simpl$NAME))    # Set a default country color
col[wrld_simpl$NAME=="Indonesia"] <- "#FF0000"    # Highlight country

plot(wrld_simpl, col=col, bg=bgcolor, border="green", ann=FALSE,  axes=FALSE, 
     xpd=FALSE,  xlim=c(-180,180), ylim=c(-90,90),  setParUsrBB=TRUE)

graphics.off()

globejs(img=earth, bg="black", lat=globe$latitude, long=globe$longitude, value=globe$Steaming.Recode, color=col, pointsize=0.6, rotationlat=0.2, rotationlong=-3.5, fov=18, atmosphere=TRUE)
```
Nb. Country borders highlighted green; Indonesia highlighted red; Arc length indicates vessel status (moving to stationary)

H1: Stationary ships are more likely to be attacked than moving ships in Southeast Asia.
H0: Stationary ships are no less, or less likely, to be attacked than moving ships in Southeast Asia.

In this globe we take a regional focus for Southeast Asia. We can observe that whether a ship was moving (dot) or stationary (arc) appears to make little difference to the number of relative pirate attacks. This could be potentially due to the archipelago maritime geography; with many ports concentrated in a small area pirates may have similar accessibility to pirating trade both for anchored ships as well as for ships underway.


### Regional Focus: Piracy in the Gulf of Aden & Arabian Sea
```{r, echo=FALSE}
######################################
# Create a Globe Viewer 3: Using 'Steaming.Recode' for arcs; Africa region focus; Somalia highlighted -LH
######################################

bgcolor <- "#000025"
earth <- tempfile(fileext=".jpg")
jpeg(earth, width=2048, height=1024, quality=100, bg=bgcolor, antialias="default")
par(mar = c(0,0,0,0), pin = c(4,2), pty = "m", xaxs = "i",
    xaxt = "n", xpd = FALSE, yaxs = "i", yaxt = "n")

col <- rep("black",length(wrld_simpl$NAME))    # Set a default country color
col[wrld_simpl$NAME=="Somalia"] <- "#FF0000"    # Highlight country

plot(wrld_simpl, col=col, bg=bgcolor, border="green", ann=FALSE,  axes=FALSE, 
     xpd=FALSE,  xlim=c(-180,180), ylim=c(-90,90),  setParUsrBB=TRUE)

graphics.off()

globejs(img=earth, bg="black", lat=globe$latitude, long=globe$longitude, value=globe$Steaming.Recode, color=col, pointsize=0.6, rotationlat=0.1, rotationlong=-2.5, fov=15, atmosphere=TRUE)
```
Nb. Country borders highlighted green; Somalia highlighted red; Arc length indicates vessel status (stationary to moving)

H1: Stationary ships are more likely to be attacked than moving ships in the Arabian Sea.
H0: Stationary ships are no less, or less likely, to be attacked than moving ships in the Arabian Sea. 

In this globe we take a regional focus for the Arabian Sea that includes the Gulf of Aden. We can observe that whether a ship was moving (dot) or stationary (arc) appears to make a large difference to the number of relative pirate attacks. Ships that were underway were more likely to be attacked. This is likely due to how large oceangoing ship traffic often transits through the Arabian Sea, but does not anchor. Pirates are therefore forced to conduct mobile pirate raids. 

#######################
#Descriptive Statistics
#######################

#Histogram of attack frequnecy
hist(shipping$`attacks/Year`)
```

Similarily, the sucess of attacks is distributed in a very similar fashion to the number of attacks.  However, our study seeks to see if there is indeed a statistically significant relationship between the number of attacks and attack success.   

```{r, echo=FALSE}
#Histogram of successful attacks per year
hist(shipping$`successful Attacks/Year`)
```

Below, a plot displays the relationship between attack success and the GDP per capita of the closest coastal state.  

```{r, echo=FALSE}
#histogram of attack success over GDP per cap
plot(shipping$`GDP per cap`, shipping$`success Ratio`)
```

Here we have seperate coplots depicting the success ratio for attacks by year, sorted by the closest coastal state. 

```{r, echo=FALSE}
#exploring the data 
coplot(shipping$`success Ratio` ~ year|country, type="l", data=shipping) # Lines
```

When we examine the heterogeneity across countries it appears that a pirates change of a successful attack is dependent on the closest coastal state.  For instance, the changes of a successful attack in the Philippines is drastically lower than the odds of a successful attack in Bangladesh.  This also holds true when considering confidence intervals which are also depicted.  

```{r, echo=FALSE}
# Fixed effects: Heterogeneity across countries (or entities)
# interestingly, attacks in Philipines have the lowest success ratio and attacks in Brazil have the highest.
# plotmeans draw a 95% confidence interval around the means
plotmeans(shipping$`success Ratio` ~ country, main="Heterogeineity across countries", data=shipping)
```

When we examine the heterogeneity across years the means seem to stay within the .65 - .85 range.  However, due to extremely large confidence intervals, time does not appear to be statistically significant.  

```{r, echo=FALSE}
# Fixed effects: Heterogeneity across years
# plotmeans draw a 95% confidence interval around the means
plotmeans(shipping$`success Ratio` ~ year, main="Heterogeineity across years", data=shipping)
```

Below is our first OLS regrssion.  Although OLS regrssion does not consider heterogeneity across groups or time, an OLS regrssion can still prove useful for gathering initial insight into the relationship of our variables.

For instance, here we see that attacks per year does not a statistically significant effect with a p-value of >.1.  However, an interesting point worth noting would be that negative coefficient of attacks per year. 

```{r}
# OLS Regression
# Regular OLS regression does not consider heterogeneity across groups or time.
# In this simple model, the number of attacks has a slightly negative relationship with attack success, however it is not stat. sig. 
ols <-lm(shipping$`success Ratio` ~ shipping$`attacks/Year`, data=shipping)
summary(ols)
```
The below plot shows that after attacks in a certain country reach a threshold, approximately 40, their attack success ratio is steadily above .6.
```{r, echo=FALSE}
# The below plot shows that after attacks in a certain country reach a threshold, approximately 40, their attack success ratio is steadily above .6
yhat <- ols$fitted
plot(shipping$`attacks/Year`, shipping$`success Ratio`, pch=19, xlab="x1", ylab="y")
abline(lm(shipping$`success Ratio`~shipping$`attacks/Year`),lwd=3, col="red")
```

In our model's first fixed effects regression, attacks per year becomes statistically significant and has a small negative coefficient of -.002.  It is important to note that the attacks per year became significant only in the fixed effects model, as opposed to the OLS.

```{r, echo=FALSE}
#attempting fixed effects model - attack success on attacks per year
fixed <- plm(shipping$`success Ratio` ~ shipping$`attacks/Year`, data=shipping, index=c("country", "year"), model="within")
summary(fixed)
```

However, once we add a variable controlling for the coast ratio of the closest coastal state, attacks per year once again becomes statistically insignificant.  

```{r, echo=FALSE}
# Increasing the complexity of our model
############################

# OLS Regression
# Regular OLS regression does not consider heterogeneity across groups or time.
# In this simple model, the number of attacks has a slightly negative relationship with attack success, however it is not stat. sig. 
# Model 2 - Success ratio on attacks/year + coast
ols2 <-lm(shipping$`success Ratio` ~ shipping$`attacks/Year` + shipping$`coast/Area ratio (m/km2)`, data=shipping)
summary(ols2)
```

Interestingly, the plot below shows that the coastal ratio of the closest coastal state most likely does have a positive, significant effect on attack success, however the Philippines is a strong outlier. 

```{r, echo=FALSE}
yhat2 <- ols$fitted
plot(shipping$`coast/Area ratio (m/km2)`, shipping$`success Ratio`, pch=19, xlab="x1", ylab="y")
abline(lm(shipping$`success Ratio`~shipping$`coast/Area ratio (m/km2)`),lwd=3, col="red")
```

When we add GDP per capita nothing is significant. 

```{r, echo=FALSE}
### Model 3 - Sucess rati on a
# attacks/year + coast + GDP
ols3 <-lm(shipping$`success Ratio` ~ shipping$`attacks/Year` + shipping$`coast/Area ratio (m/km2)` + shipping$`GDP per cap`, data=shipping)
summary(ols3)
```

A graph of the above regression

```{r, echo=FALSE}
yhat3 <- ols$fitted 
plot(shipping$`GDP per cap`, shipping$`success Ratio`, pch=19, xlab="x1", ylab="y")
abline(lm(shipping$`success Ratio`~shipping$`GDP per cap`),lwd=3, col="red")
```

Number of attacks is significant again.
```{r, echo=FALSE}
# FE model 2, Sucess Ratio on Attacks/Year + GDP per cap 
#attempting fixed effects model
fixed2 <- plm(shipping$`success Ratio` ~ shipping$`attacks/Year` + shipping$`GDP per cap`, data=shipping, index=c("country", "year"), model="within")
summary(fixed2)
```

Here are diagnostic regressions

```{r}
#####################
# Regression Diagnostics
######################

#Assessing homoscedasticity (we have met the constant variance assumption if p < 1.95)
ncvTest(ols2)
```

```{r}
#Assessing multicollinearity
vif(ols2)
```

```{r}
#Assessing outliers
outlierTest(ols2)
```



# V Insights & Further Development

* Attacks per year for the nation-states selected occurred mostly within the 0-20 range. 
* Attacks in Bangladesh were statisticially far more likely to succeed than an attack on a shipping vessel in the Philippines, even accounting for 5% confidence intervals. This would suggest that Baglandeshi maritime patrols are inferior to those currently in place in the Philippines. 
* Considering the progress of time for piracy attacks, indicated by our data is that attackers have not grown more successful. This is evidenced by the very large confidence intervals observable in the 'Heterogeneity across Years' table. 
* There exists a small and negative linear coefficient for the amount of attacks per year correlated with success of these attacks, that may suggest that maritime patrol is improving in effectiviness over time. 
* The Philippines occurs as a strong outlier when considering our 'Coast-Line Ratio' control for attacks per country (most likely given the archipelago geography). A further point for development of our model would be to remove the Philippines as a strong outlier when regressing coast-line ratio with success of an attack.
