---
title: "Determinants of Maritime Piracy"
author: "Bjoern Boening, Laurence Hendry & Cody Koebnick"
date: "4 December 2015"
output:
  slidy_presentation:
   font_adjustment: -0.5
  ioslides_presentation:
    transition: slower
    logo: logo.png
---

```{r, echo=FALSE, message=FALSE}
#helpful shortcuts for editing what is includeded in Knitr: #http://kbroman.org/knitr_knutshell/pages/Rmarkdown.html
library(base)
library(rio) # swiss army knife for imports
library(plyr) # count occurences
library(dplyr) # data wrangling
library(tidyr) # data wrangling
library(ggplot2) # nice plots
library(stargazer) # nicer regression output which looks like a real publication
library(car) # scatterplots 
library(httr) # scraping from http sites
library(XML) # Tool for generating XML file
library(WDI) # Scraping Data from the World Bank
library(countrycode) # provides world bank country codes 
library(gplots)
library(plm)
library(knitr)
library(threejs) # makes globes
library(maptools)
library(maps)
library(reshape)
library(pander)
library(Amelia)

try(setwd("E:/bjoer/Documents/GitHub/MaritimePiracy"), silent=TRUE)
try(setwd("/Users/laurencehendry/GitHub/MaritimePiracy"),silent=TRUE) 
try(setwd("/Users/codykoebnick/Documents/MaritimePiracy"),silent=TRUE)
shipping <- read.csv("shippingraw.csv", header = TRUE, sep = ",", stringsAsFactors = FALSE, na.strings = c("", "NA"))

#renaming and dropping some of our columns 
names(shipping)[1] <- 'country'
names(shipping)[2] <- 'year'
shipping$X1 = NULL
names(shipping)[3] <- 'coast/Area ratio (m/km2)'
names(shipping)[4] <- 'GDP per cap'
names(shipping)[5] <- 'attacks/Year'
names(shipping)[6] <- 'successful Attacks/Year'
names(shipping)[7] <- 'success Ratio'
```

## Contents
- Introduction & Contents
    + Maritime Piracy:  Our Data for this 21st Century Problem
- Laurence
    + Global Overview: Piracy and Maritime Defence Propensity 
    + Regional Focus: Piracy in the South China Sea
    + Regional Focus: Piracy in the Gulf of Aden & Arabian Sea
- Cody
    + Aggregated Data: The 8 Countries with Highest Amounts of Piracy  
    + Adding Variables:  Military Expenditure and Military Conflicts
    + Regression Results 
- Bjoern
    + Model
    + Model & Chi Square Test
    + Innocent Logistic Model
    + Scrutinise Us

## Global Overview: Piracy and Maritime Defence Propensity {.smaller .build}

```{r, echo=FALSE}
try(setwd("/Users/laurencehendry/GitHub/MaritimePiracy"),silent=TRUE) 
try(setwd("E:/bjoer/Documents/GitHub/MaritimePiracy"), silent=TRUE)
globe <- read.csv("MaritimePiracyTennessee.csv", header = TRUE, sep = ";", stringsAsFactors = FALSE, na.strings = c("", "NA"))

URL <- 'https://en.wikipedia.org/wiki/List_of_countries_by_length_of_coastline'

tables <- URL %>% GET() %>%
  content(as = 'parsed') %>% 
  readHTMLTable()

#names(tables)

CoastlineTable <- tables[[1]]

#head(CoastlineTable)[, 1:3]

#cleaning the scrape for merge
CoastlineTable$V2 = NULL
CoastlineTable$V3 = NULL
CoastlineTable$V4 = NULL
CoastlineTable$V5 = NULL
CoastlineTable$V6 = NULL
CoastlineTable$V8 = NULL

#renaming, part of scrape
colnames(CoastlineTable)
#names(CoastlineTable)
names(CoastlineTable)[1] <- 'closest_coastal_state'
names(CoastlineTable)[2] <- 'CoastRatio'

#merging
#p297 from R for Dummies
globe <- merge(globe, CoastlineTable, all.x=TRUE)


globe$longitude <- gsub(",", ".", globe$longitude)
globe$latitude <- gsub(",", ".", globe$latitude)

data(wrld_simpl, package="maptools")
data(worldMapEnv, package="maps")
```

```{r, echo=FALSE}

######################################
# Create a Globe Viewer: Using 'CoastRatio' for arcs; zoomed out region focus; Phillipines highlighted -LH 
######################################

bgcolor <- "#000025"
earth <- tempfile(fileext=".jpg")
jpeg(earth, width=2048, height=1024, quality=100, bg=bgcolor, antialias="default")
par(mar = c(0,0,0,0), pin = c(4,2), pty = "m", xaxs = "i",
    xaxt = "n", xpd = FALSE, yaxs = "i", yaxt = "n")

col <- rep("black",length(wrld_simpl$NAME))    # Set a default country color
col[wrld_simpl$NAME=="Philippines"] <- "#FF0000"    # Highlight country


plot(wrld_simpl, col=col,          bg=bgcolor, border="green", ann=FALSE,  axes=FALSE, 
     xpd=FALSE,  xlim=c(-180,180), ylim=c(-90,90),  setParUsrBB=TRUE)

graphics.off()

globejs(img=earth, bg="black", lat=globe$latitude, long=globe$longitude, value=globe$CoastRatio, color=col, pointsize=0.5, rotationlat=0.2,     rotationlong=-3, fov=30, atmosphere=TRUE)
```

- Country borders highlighted green; Philippines highlighted red; Arc length indicates height of coast-to-country ratio (small to large)
- ***H1: Most pirate attacks occur in areas where countries have smaller country-to-coast ratios, and thus are less incentivised to defend against maritime pirate  attacks.***
- H0: Whether a country has a greater country-to-coast ratio has either an inverse or no effect on where pirate attacks occur.
- This is not always the case, depending on region. Our H1 remains true for the area around the Gulf of Aden and particularly in the Arabian Sea, but not for the South China Sea. 
- 3D Globes using the 'threejs' Package: The [3D Globe package](file:///Users/laurencehendry/GoogleDrive/Master%20Thesis%20-%20Shared/MPP-E1180%20-%20Introduction%20to%20Collaborative%20Social%20Science%20Data%20Analysis/Laurence-Github/Working%203D%20Globe/Model%20v1.html)


## Regional Focus: Piracy in the South China Sea {.smaller}
```{r, echo=FALSE}
######################################
# Create a Globe Viewer 2: Using 'Steaming.Recode' for arcs; SEAsia region focus; Indonesia highlighted -LH
######################################

bgcolor <- "#000025"
earth <- tempfile(fileext=".jpg")
jpeg(earth, width=2048, height=1024, quality=100, bg=bgcolor, antialias="default")
par(mar = c(0,0,0,0), pin = c(4,2), pty = "m", xaxs = "i",
    xaxt = "n", xpd = FALSE, yaxs = "i", yaxt = "n")

col <- rep("black",length(wrld_simpl$NAME))    # Set a default country color
col[wrld_simpl$NAME=="Indonesia"] <- "#FF0000"    # Highlight country

plot(wrld_simpl, col=col, bg=bgcolor, border="green", ann=FALSE,  axes=FALSE, 
     xpd=FALSE,  xlim=c(-180,180), ylim=c(-90,90),  setParUsrBB=TRUE)

graphics.off()

globejs(img=earth, bg="black", lat=globe$latitude, long=globe$longitude, value=globe$Steaming.Recode, color=col, pointsize=0.6, rotationlat=0.2, rotationlong=-3.5, fov=18, atmosphere=TRUE)
```

- Country borders highlighted green; Indonesia highlighted red; Arc length indicates vessel status (moving to stationary)
- ***H1: Stationary ships are more likely to be attacked than moving ships in Southeast Asia.***
- H0: Stationary ships are no less, or less likely, to be attacked than moving ships in Southeast Asia.
- We can observe that whether a ship was moving (dot) or stationary (arc) appears to make little difference to the number of relative pirate attacks. 

## Regional Focus: Piracy in the Gulf of Aden & Arabian Sea {.smaller}
```{r, echo=FALSE}
######################################
# Create a Globe Viewer 3: Using 'Steaming.Recode' for arcs; Africa region focus; Somalia highlighted -LH
######################################

bgcolor <- "#000025"
earth <- tempfile(fileext=".jpg")
jpeg(earth, width=2048, height=1024, quality=100, bg=bgcolor, antialias="default")
par(mar = c(0,0,0,0), pin = c(4,2), pty = "m", xaxs = "i",
    xaxt = "n", xpd = FALSE, yaxs = "i", yaxt = "n")

col <- rep("black",length(wrld_simpl$NAME))    # Set a default country color
col[wrld_simpl$NAME=="Somalia"] <- "#FF0000"    # Highlight country

plot(wrld_simpl, col=col, bg=bgcolor, border="green", ann=FALSE,  axes=FALSE, 
     xpd=FALSE,  xlim=c(-180,180), ylim=c(-90,90),  setParUsrBB=TRUE)

graphics.off()

globejs(img=earth, bg="black", lat=globe$latitude, long=globe$longitude, value=globe$Steaming.Recode, color=col, pointsize=0.6, rotationlat=0.1, rotationlong=-2.5, fov=15, atmosphere=TRUE)
```

- Country borders highlighted green; Somalia highlighted red; Arc length indicates vessel status (stationary to moving)
- ***H1: Stationary ships are more likely to be attacked than moving ships in the Arabian Sea.***
- H0: Stationary ships are no less, or less likely, to be attacked than moving ships in the Arabian Sea. 
- We can observe that whether a ship was moving (dot) or stationary (arc) appears to make a large difference to the number of relative pirate attacks. Ships that were underway were more likely to be attacked. 


## Aggregated Data: The 8 Countries with Highest Amounts of Piracy 
```{r, echo=FALSE}
# import data
# empty cells are now coded with NA and can manually be excluded from any function with na.omit command
#shipping <- read.csv("shippingraw.csv", header = TRUE, sep = ",", stringsAsFactors = FALSE, na.strings = c("", "NA"))
# have a look at how the variables are created
#str(shipping)

#renaming and dropping some of our columns 
#names(shipping)[1] <- 'country'
#names(shipping)[2] <- 'year'
#shipping$X1 = NULL
#names(shipping)[3] <- 'coast/Area ratio (m/km2)'
#names(shipping)[4] <- 'GDP per cap'
#names(shipping)[5] <- 'attacks/Year'
#names(shipping)[6] <- 'successful Attacks/Year'
#names(shipping)[7] <- 'success Ratio'
# These are the two graphs to show on this slide 
plotmeans(shipping$`success Ratio` ~ country, main="Average Success Ratio of Pirates per Country", data=shipping, xlab="Country", ylab="Success Ratio")

plotmeans(shipping$`success Ratio` ~ year, main="Average Success Ratio of Pirates per Year", data=shipping, xlab="Year", ylab="Success Ratio")
```

## Adding Variables:  Military Expenditure and Military Conflicts 
```{r, echo=FALSE}
getwd()
try(setwd("/Users/laurencehendry/GitHub/MaritimePiracy"), silent=TRUE) 
#try(setwd("E:/bjoer/Documents/GitHub/MaritimePiracy"), silent=TRUE)
#import data
# empty cells are now coded with NA and can manually be excluded from any function with na.omit command
ucdp.prio <- read.csv("124920_1ucdpprio-armed-conflict-dataset_v.4-2015.csv", header = TRUE, sep = ",", stringsAsFactors = FALSE, na.strings = c("", "NA"))
# have a look at how the variables are created

names(ucdp.prio)[3] <- 'country'
names(ucdp.prio)[2] <- 'year'

ucdp.prio$ConflictId <- NULL
ucdp.prio$SideA <- NULL
ucdp.prio$SideA2nd <- NULL
ucdp.prio$SideB <- NULL
ucdp.prio$SideBID <- NULL
ucdp.prio$SideB2nd <- NULL
ucdp.prio$Incompatibility <- NULL
ucdp.prio$TerritoryName <- NULL
ucdp.prio$CumulativeIntensity <- NULL
ucdp.prio$TypeOfConflict <- NULL
ucdp.prio$StartDate <- NULL
ucdp.prio$StartPrec <- NULL
ucdp.prio$StartDate2 <- NULL
ucdp.prio$StartPrec2 <- NULL
ucdp.prio$EpEnd <- NULL
ucdp.prio$EpEndDate <- NULL
ucdp.prio$EpEndPrec <- NULL
ucdp.prio$GWNoA <- NULL
ucdp.prio$GWNoA2nd <- NULL
ucdp.prio$GWNoB <- NULL
ucdp.prio$GWNoB2nd <- NULL
ucdp.prio$GWNoLoc <- NULL
ucdp.prio$Region <- NULL
ucdp.prio$Version <- NULL
#str(shipping)

ucdp.prio <- ddply(ucdp.prio, .(country, year), numcolwise(sum))

ship.ucdp <- merge(shipping, ucdp.prio, by=c("country","year"), all.x=TRUE) 
remove(list=c("shipping"))
remove(list=c("ucdp.prio"))

ship.ucdp <- unique(ship.ucdp[ , 1:8 ] )

ship.ucdp$IntensityLevel[is.na(ship.ucdp$IntensityLevel)] <- 0 

#######################
#adding military expenditure numbers
######################
#import data
# empty cells are now coded with NA and can manually be excluded from any function with na.omit command
Military <- read.csv("Military.csv", header = TRUE, sep = ",", stringsAsFactors = FALSE, na.strings = c("", "NA"))
# have a look at how the variables are created

names(Military)[1] <- 'country'
names(Military)[2] <- '1992'
names(Military)[3] <- '1993'
names(Military)[4] <- '1994'
names(Military)[5] <- '1995'
names(Military)[6] <- '1996'
names(Military)[7] <- '1997'
names(Military)[8] <- '1998'
names(Military)[9] <- '1999'
names(Military)[10] <- '2000'
names(Military)[11] <- '2001'
names(Military)[12] <- '2002'
names(Military)[13] <- '2003'
names(Military)[14] <- '2004'
names(Military)[15] <- '2005'
names(Military)[16] <- '2006'
names(Military)[17] <- '2007'
names(Military)[18] <- '2008'
names(Military)[19] <- '2009'
names(Military)[20] <- '2010'
names(Military)[21] <- '2011'
names(Military)[22] <- '2012'
names(Military)[23] <- '2013'
names(Military)[24] <- '2014'

x2 <- melt(Military,id=c("country"),variable_name="Year")
x2[,"Year"] <- as.numeric(gsub("X","",x2[,"Year"]))

names(x2)[2] <- 'year'
names(x2)[3] <- 'Military Expenditure'

ship.ucdp.mil <- merge(ship.ucdp, x2, by=c("country","year"), all.x=TRUE) 
remove(list=c("Military"))
remove(list=c("x2"))
remove(list=c("ship.ucdp"))

#histograms to display on slide
plot(ship.ucdp.mil$IntensityLevel, ship.ucdp.mil$`Military Expenditure`, xlab="Intensity Level", ylab="Military Expenditure")

plot(ship.ucdp.mil$`Military Expenditure`, ship.ucdp.mil$`attacks/Year`, xlab="Military Expenditure", ylab="Attacks Per Year")
```

## Regression Results {.smaller}

```{r, echo=FALSE}
###############
## Inferential Statistics
###############
### I would like these two regressions to display
ols10 <-lm(ship.ucdp.mil$`attacks/Year` ~ ship.ucdp.mil$`coast/Area ratio (m/km2)` + ship.ucdp.mil$`GDP per cap` + ship.ucdp.mil$`Military Expenditure`, data=ship.ucdp.mil)
pander(ols10)

fixed8 <- plm(ship.ucdp.mil$`attacks/Year` ~ ship.ucdp.mil$`Military Expenditure` + ship.ucdp.mil$`GDP per cap`, data=ship.ucdp.mil, index=c("country", "year"), model="within",)
summary(fixed8)
```

```{r, echo=FALSE}
######
#IMPORT & DATA WRANGLING
######
# set working directories 
try(setwd("/Users/codykoebnick/Documents/MaritimePiracy"))
try(setwd("E:/bjoer/Documents/GitHub/MaritimePiracy"))
try(setwd("/Users/laurencehendry/GitHub/MaritimePiracy")) 
#getwd()
#import data
# empty cells are coded with NA and can manually be excluded from any function with na.omit command
micro <- read.csv("MaritimePiracyTennessee.csv", header = TRUE, sep = ";", stringsAsFactors = FALSE, na.strings = c("", "NA"))
# have a look at how the variables are created
#str(micro)
#subsetting (keep) variables
sub <- micro[c(4, 6, 12, 18, 24, 23)]

#renaming and recoding
names(sub)[1] <- 'year'

names(sub)[2] <- 'time'
sub$time <- factor(sub$time,
                    levels = c(1,2,3,4))

names(sub)[3] <- 'state'

names(sub)[4] <- 'type'
sub$type[sub$type==1] <- 222
sub$type[sub$type==5] <- 222
sub$type[sub$type==9] <- 222
sub$type[sub$type==2] <- 111
sub$type[sub$type==3] <- 111
sub$type[sub$type==4] <- 111
sub$type[sub$type==6] <- 111
sub$type[sub$type==7] <- 111
sub$type[sub$type==8] <- 111
sub$type[sub$type==111] <- 1
sub$type[sub$type==222] <- 2
sub$type <- factor(sub$type)
names(sub)[5] <- 'incident'
sub$incident <- factor(sub$incident)
names(sub)[6] <- 'stat'
sub$stat <- factor(sub$stat,
                    levels = c(1,2,3,4))
sub$status <- recode(sub$stat, "c(1)='1'; c(2,3,4)='2'") # what a bastard this line was arrgg
sub$stat = NULL

# Delete missing values
#table(sub$year, useNA = "always")
#table(sub$time, useNA = "always")
sub$time[sub$time==-99] <- NA
#table(sub$status, useNA = "always")
sub$status[sub$status==-99] <- NA
#table(sub$type, useNA = "always")
sub$type[sub$type==-99] <- NA
sub$type[sub$type==22] <- NA
sub$type[sub$type==696] <- NA
sub$type[sub$type==10] <- NA
#table(sub$incident, useNA = "always")
sub$incident[sub$incident==-99] <- NA
#table(sub$state, useNA = "always")
# Omit NAs # not working must create a new data frame
#sub$time <- na.omit(sub$time)
#sub$status <- na.omit(sub$status)
#sub$type <- na.omit(sub$type)
#sub$incident <- na.omit(sub$incident)

```


## Model

What has determined maritime piracy between 1995 and 2014?

```{r, echo=FALSE}
suc_ratio <- ggplot(na.omit(sub))
suc_ratio + aes(factor(incident)) + geom_bar()
suc_ratio <- geom_bar()

```

## Model & Chi Square Test

- H1. The success rate of an attacks shrinks for big ships,
- H2. And for vessels in motion,
- H3. And during the day.

```{r, echo=FALSE}
sub$incident <- factor(sub$incident)
sub$type <- factor(sub$type)
sub$status <- factor(sub$status)
sub$time <- factor(sub$time)
tab <- table(na.omit(sub)$incident,na.omit(sub)$type) # A will be rows, B will be columns
tab1 <- table(na.omit(sub)$incident,na.omit(sub)$status) # A will be rows, B will be columns 
tab2 <- table(na.omit(sub)$incident,na.omit(sub)$time) # A will be rows, B will be columns 
```

### Ship Type
```{r, echo=FALSE}
summary(tab, digits = 2)
```
### Ship status (Factor)
```{r, echo=FALSE}
summary(tab1, digits = 2)
```
### Time of day (Factor)
```{r, echo=FALSE}
summary(tab2, digits = 2)
```


## Innocent Logistic Model

```{r, echo=FALSE, results="asis"}
model <- glm(incident ~ type,  family=binomial(link='logit'),data=sub)
#summary(model)
stargazer(model, digits = 2,  title="Regression Results",  type = "html")
```


## Srutinize us

```{r, echo=FALSE}
missmap(sub, main = "Too much NAs?")
```
