---
title: "Determinants of Maritime Piracy" 
author: "Bjoern Boening, Laurence Hendry & Cody Koebnick"
date: "4 December 2015"
output: 
  slidy_presentation:
    font_adjustment: -1
---

```{r, echo=FALSE, message=FALSE}
#helpful shortcuts for editing what is includeded in Knitr: http://kbroman.org/knitr_knutshell/pages/Rmarkdown.html
library(base)
library(rio) # swiss army knife for imports
library(plyr) # count occurences
library(dplyr) # data wrangling
library(tidyr) # data wrangling
library(ggplot2) # nice plots
library(stargazer) # nicer regression output which looks like a real publication
library(car) # scatterplots 
library(httr) # scraping from http sites
library(XML) # Tool for generating XML file
library(WDI) # Scraping Data from the World Bank
library(countrycode) # provides world bank country codes 
library(gplots)
library(plm)
library(knitr)
library(threejs)
library(maptools)
library(maps)


try()
try(setwd("/Users/laurencehendry/GitHub/MaritimePiracy"),silent=TRUE) 
try(setwd("/Users/codykoebnick/Documents/MaritimePiracy"),silent=TRUE)
shipping <- read.csv("shippingraw.csv", header = TRUE, sep = ",", stringsAsFactors = FALSE, na.strings = c("", "NA"))

#renaming and dropping some of our columns 
names(shipping)[1] <- 'country'
names(shipping)[2] <- 'year'
shipping$X1 = NULL
names(shipping)[3] <- 'coast/Area ratio (m/km2)'
names(shipping)[4] <- 'GDP per cap'
names(shipping)[5] <- 'attacks/Year'
names(shipping)[6] <- 'successful Attacks/Year'
names(shipping)[7] <- 'success Ratio'
```

## Contents
- Introduction
+ A 21st Century Problem: Maritime Piracy
+ Our Data
- Bjoern
+ Model
+ Analysis
+ Results
- Laurence
+ 3D Globe
+ 3D Region: Southeast Asia
+ 3D Region: Arabian Sea
- Cody
+ Model 
+ Analysis
+ Results

## Visualizing and Examining the Data

3D Globes using the 'threejs' Package: The [threejs package](http://bwlewis.github.io/rthreejs/). The 3D globes are full navigable, with pinch-zooming enabled for trackpads to help with examining statistical results in particular regions. 

```{r, echo=FALSE}
try(setwd("/Users/laurencehendry/GitHub/MaritimePiracy"),silent=TRUE) 
globe <- read.csv("MaritimePiracyTennessee.csv", header = TRUE, sep = ";", stringsAsFactors = FALSE, na.strings = c("", "NA"))

URL <- 'https://en.wikipedia.org/wiki/List_of_countries_by_length_of_coastline'

tables <- URL %>% GET() %>%
  content(as = 'parsed') %>% 
  readHTMLTable()

names(tables)

CoastlineTable <- tables[[1]]

head(CoastlineTable)[, 1:3]

#cleaning the scrape for merge
CoastlineTable$V2 = NULL
CoastlineTable$V3 = NULL
CoastlineTable$V4 = NULL
CoastlineTable$V5 = NULL
CoastlineTable$V6 = NULL
CoastlineTable$V8 = NULL

#renaming, part of scrape
colnames(CoastlineTable)
names(CoastlineTable)
names(CoastlineTable)[1] <- 'closest_coastal_state'
names(CoastlineTable)[2] <- 'CoastRatio'

#merging
#p297 from R for Dummies
globe <- merge(globe, CoastlineTable, all.x=TRUE)


globe$longitude <- gsub(",", ".", globe$longitude)
globe$latitude <- gsub(",", ".", globe$latitude)

data(wrld_simpl, package="maptools")
data(worldMapEnv, package="maps")
```

## Global Overview: Piracy and Maritime Defence Propensity 
```{r, echo=FALSE}

######################################
# Create a Globe Viewer: Using 'CoastRatio' for arcs; zoomed out region focus; Phillipines highlighted -LH 
######################################

bgcolor <- "#000025"
earth <- tempfile(fileext=".jpg")
jpeg(earth, width=2048, height=1024, quality=100, bg=bgcolor, antialias="default")
par(mar = c(0,0,0,0), pin = c(4,2), pty = "m", xaxs = "i",
    xaxt = "n", xpd = FALSE, yaxs = "i", yaxt = "n")

col <- rep("black",length(wrld_simpl$NAME))    # Set a default country color
col[wrld_simpl$NAME=="Philippines"] <- "#FF0000"    # Highlight country


plot(wrld_simpl, col=col,          bg=bgcolor, border="green", ann=FALSE,  axes=FALSE, 
     xpd=FALSE,  xlim=c(-180,180), ylim=c(-90,90),  setParUsrBB=TRUE)

graphics.off()

globejs(img=earth, bg="black", lat=globe$latitude, long=globe$longitude, value=globe$CoastRatio, color=col, pointsize=0.5, rotationlat=0.2,     rotationlong=-3, fov=30, atmosphere=TRUE)
```
- Country borders highlighted green; Philippines highlighted red; Arc length indicates height of coast-to-country ratio (small to large)
***H1: Most pirate attacks occur in areas where countries have smaller country-to-coast ratios, and thus are less incentivised to defend against maritime pirate  attacks. 
***H0: Whether a country has a greater country-to-coast ratio has either an inverse or no effect on where pirate attacks occur. 
- This is not always the case, depending on region. Our H1 remains true for the area around the Gulf of Aden and particularly in the Arabian Sea, but not for the South China Sea. 

## Regional Focus: Piracy in the South China Sea
```{r, echo=FALSE}
######################################
# Create a Globe Viewer 2: Using 'Steaming.Recode' for arcs; SEAsia region focus; Indonesia highlighted -LH
######################################

bgcolor <- "#000025"
earth <- tempfile(fileext=".jpg")
jpeg(earth, width=2048, height=1024, quality=100, bg=bgcolor, antialias="default")
par(mar = c(0,0,0,0), pin = c(4,2), pty = "m", xaxs = "i",
    xaxt = "n", xpd = FALSE, yaxs = "i", yaxt = "n")

col <- rep("black",length(wrld_simpl$NAME))    # Set a default country color
col[wrld_simpl$NAME=="Indonesia"] <- "#FF0000"    # Highlight country

plot(wrld_simpl, col=col, bg=bgcolor, border="green", ann=FALSE,  axes=FALSE, 
     xpd=FALSE,  xlim=c(-180,180), ylim=c(-90,90),  setParUsrBB=TRUE)

graphics.off()

globejs(img=earth, bg="black", lat=globe$latitude, long=globe$longitude, value=globe$Steaming.Recode, color=col, pointsize=0.6, rotationlat=0.2, rotationlong=-3.5, fov=18, atmosphere=TRUE)
```
- Country borders highlighted green; Indonesia highlighted red; Arc length indicates vessel status (moving to stationary)
***H1: Stationary ships are more likely to be attacked than moving ships in Southeast Asia.
***H0: Stationary ships are no less, or less likely, to be attacked than moving ships in Southeast Asia.
- We can observe that whether a ship was moving (dot) or stationary (arc) appears to make little difference to the number of relative pirate attacks. 

## Regional Focus: Piracy in the Gulf of Aden & Arabian Sea
```{r, echo=FALSE}
######################################
# Create a Globe Viewer 3: Using 'Steaming.Recode' for arcs; Africa region focus; Somalia highlighted -LH
######################################

bgcolor <- "#000025"
earth <- tempfile(fileext=".jpg")
jpeg(earth, width=2048, height=1024, quality=100, bg=bgcolor, antialias="default")
par(mar = c(0,0,0,0), pin = c(4,2), pty = "m", xaxs = "i",
    xaxt = "n", xpd = FALSE, yaxs = "i", yaxt = "n")

col <- rep("black",length(wrld_simpl$NAME))    # Set a default country color
col[wrld_simpl$NAME=="Somalia"] <- "#FF0000"    # Highlight country

plot(wrld_simpl, col=col, bg=bgcolor, border="green", ann=FALSE,  axes=FALSE, 
     xpd=FALSE,  xlim=c(-180,180), ylim=c(-90,90),  setParUsrBB=TRUE)

graphics.off()

globejs(img=earth, bg="black", lat=globe$latitude, long=globe$longitude, value=globe$Steaming.Recode, color=col, pointsize=0.6, rotationlat=0.1, rotationlong=-2.5, fov=15, atmosphere=TRUE)
```
- Country borders highlighted green; Somalia highlighted red; Arc length indicates vessel status (stationary to moving)
***H1: Stationary ships are more likely to be attacked than moving ships in the Arabian Sea.
***H0: Stationary ships are no less, or less likely, to be attacked than moving ships in the Arabian Sea. 
- We can observe that whether a ship was moving (dot) or stationary (arc) appears to make a large difference to the number of relative pirate attacks. Ships that were underway were more likely to be attacked. 

